---
title: "TBT_DMSO_Final.Rmd"
author: "Taylor"
date: '2022-06-23'
output: html_document
---

#1vs1 TBT DMSO
This is the RMArkdown for my TBT vs DMSO data. I will be Running all of the samples together as well as subsetting by day 
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Documents/Git/MSC_Bulk/RNA_seq_Data/TBT_results/")
suppressPackageStartupMessages({
  library(ComplexHeatmap)
  library(pheatmap)
  library(RColorBrewer)
  library(DESeq2)
  library(BSgenome.Hsapiens.UCSC.hg38)
  library(tidyverse)
  library(dbplyr)
  library(annotables)
  library(org.Hs.eg.db)
  library(AnnotationDbi)
  library(TxDb.Hsapiens.UCSC.hg38.knownGene)
  library(EnsDb.Hsapiens.v86)
  library(EnhancedVolcano)
})
```
```{r load variables}
load("~/DESeq2_Variables_Preanalyzed.RData") #this will have all of the preanalyzed data
```
#Basic analysis
This starts out with the basic analysis of all the sample for with the condition(TBT,TBT,control) as the object that the program is comparing the samples with. 
This has Three Main sections:
1. Basic Analysis, for time and donciiton
2. Basic Analysis without MSC control
3. Day by day comparison
These will help to show that the most variability between the Condition and control. So I will subset the data. 


```{r, Basic Analysis  eval=FALSE}
#BiocManager::install("DESeq2")
suppressPackageStartupMessages(library("DESeq2"))

#Subsetting the data to only have the pertinent information
cts_mat_TBT<- cts_mat[,which(substr(colnames(cts_mat),2,2)=="T"|substr(colnames(cts_mat),2,2)=="D" | substr(colnames(cts_mat),1,2)=="AO")]


coldata_TBT <- coldata_norm[which(substr(rownames(coldata_norm),2,2)=="T"|substr(rownames(coldata_norm),2,2)=="D"|substr(rownames(coldata_norm),1,2)=="AO"),]
dim(cts_mat_TBT) #58037 30
dim(coldata_TBT) #31 5
coldata_TBT
##Creating DDS object from the subset data                                  

dds_TBT_TC <- DESeqDataSetFromMatrix(countData = cts_mat_TBT,
                                                 colData = coldata_TBT,
                                                 design = ~ Time + condition
                                                 )
summary(dds_TBT_TC)
head(dds_TBT_TC)   
#####Adding features and metadata columns
rownames(dds_TBT_TC) <- substr(rownames(dds_TBT_TC),1,15)
featureData <- data.frame(ensgene= substr(rownames(cts_mat_TBT),1,15))
dim(featureData)
featureData$symbol <- cts_1021$SYMBOL# Adding mcol symbol data
mcols(dds_TBT_TC) <- DataFrame(mcols(dds_TBT_TC), featureData)
mcols(dds_TBT_TC)
```

## Filtering DDS Object

```{r Filtering a dds object,eval=FALSE}
keep <- rowSums(counts(dds_TBT_TC)) >= 50 #25143 genes passed this low bar
dds_TBT_TC <- dds_TBT_TC[keep,]
dim(dds_TBT_TC)
```

Setting factor levels. It can be useful to designate factor levels while
formatting your dds object since it will influence the manner in which
DESeq compares different groups within your data.

You can specify levels either using factor()

```{r Generating dds object, eval=FALSE}
dds_TBT_TC$condition <- factor(dds_TBT_TC$condition, levels = c("control","TBT"))
dds_TBT_TC$condition
####Making things factors
dds_TBT_TC$Day <- factor(dds_TBT_TC$Day)
dds_TBT_TC$Time <- factor(dds_TBT_TC$Time)
```
##NORMALIZE: This is self-explanitory. This will normalize the data so
that we aren't getting a lot of outliers.

```{r normalize counts}
# Determine the size factors to use for normalization
dds_TBT_TC<- estimateSizeFactors(dds_TBT_TC)
sizeFactors(dds_TBT_TC)
# Extract the normalized counts
normalized_counts_TBT <- counts(dds_TBT_TC, normalized=T)
head(normalized_counts_TBT)
```
###VST vs rlog
VST is much faster to compute and less sensitive to outliers than the rlog. rlog is better for smaller data sets with samples sizes below ~30. you can perform both VST and rlog if you are unsure and then compare the meanSdPlot  or PCAs to see which generates better data. 
```{r correlation plot}
suppressPackageStartupMessages(library(pheatmap))

# Transform the normalized counts 
vsd_TBT_TC <- vst(dds_TBT_TC,blind=T )

# Extract the matrix of transformed counts
vsd_mat_TBT_TC <- assay(vsd_TBT_TC)

# Compute the correlation values between samples
vsd_cor_TBT_TC <- cor(vsd_mat_TBT_TC) 

# Plot the heatmap
pheatmap(vsd_cor_TBT_TC)
```
###PCA time
```{r pca of all RNA-seq data}
# Transform the normalized counts 
#vsd_TBT_TC <- vst(dds_TBT_TC,blind=F)

#determine what to dolor the data by 
colData(vsd_TBT_TC)
# Plot the PCA of PC1 and PC2
plotPCA(vsd_TBT_TC, intgroup= "condition")
```
##Top variance
```{r top Variance genes}
library("genefilter")
#Lets us thee variance tables to get the samples that had the biggest discrepancy for day 0 to day 14. Then create a heatmap to visualize the top genes that plat this role. We can use this to see which of the MSCs played a larger role in the definithion of MSCs by what genes 
VarGenes_TBT_TC <- head(order(rowVars(assay(vsd_TBT_TC)),decreasing = TRUE),50)
#Create a matrix of the top genes 
VG_mat_SD_NE  <- assay(vsd_SD_NoExtra)[VarGenes_SD_NE,]
head(vsd_TBT_TC)
colnames(vsd_TBT_TC)
VG_mat_SD_NE  <- VG_mat_SD_NE - rowMeans(VG_mat_SD_NE)
anno <- as.data.frame(colData(vsd_TBT_TC)[, c("condition","Time")])

library(AnnotationDbi)
library(org.Hs.eg.db)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(EnsDb.Hsapiens.v86)
edb <- EnsDb.Hsapiens.v86
organism(edb)
keytypes(edb)
keys(edb,keytype = "GENEID")

## Annotating significant table
dim(VG_mat_SD_NE)#Significant results day 0 to day 14 ordered by padj values the data rames has dimisions of 13013 6 
VG_mat_SD_NE_annoDbi <- data.frame(VG_mat_SD_NE)
VG_mat_SD_NE_annoDbi$symbol<-                                                                                      mapIds(edb,keys=rownames(VG_mat_SD_NE_annoDbi),column="SYMBOL",keytype="GENEID",multiVals="first")
dim(VG_mat_SD_NE_annoDbi) #13013 7 
rownames(VG_mat_SD_NE) <- VG_mat_SD_NE_annoDbi$symbol
pheatmap(VG_mat_SD_NE, annotation_col = anno)
```
##PCA varaition
```{r}
plotPCA(vsd_TBT_TC, intgroup = c("condition","Time"))
```
#Results
```{r Differential Expression calc, echo=FALSE, eval= FALSE}
#We didn't need to remave any of the samples so the basic analyis will be the same by conditon

DES_TBT_TC <- DESeq(dds_TBT_TC)
res_TBT_TC <- results(DES_TBT_TC)
head(res_TBT_TC)
mcols(res_TBT_TC)
```
## Filtering results
I am adding a specific contrast and alpha values to help remove some of the extra low hanging fruit.  
```{r Differential Expression results}
res_contr_TBT_TC <- results(DES_TBT_TC, contrast=c("condition","TBT","control"),alpha = 0.05)
head(res_contr_TBT_TC)
########Annotate significant genes#################
library(AnnotationDbi)
library(org.Hs.eg.db)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(EnsDb.Hsapiens.v86)
edb <- EnsDb.Hsapiens.v86
organism(edb)
keytypes(edb)
#keys(edb,keytype = "GENEID")

## Annotating analyzed results the data can be subset for analysis and significance later
dim(res_contr_TBT_TC)#Significant results day 0 to day 14 ordered by padj values the data rames has dimisions of 13013 6 
res_contr_TBT_TC_anno <- data.frame(res_contr_TBT_TC)
res_contr_TBT_TC_anno$symbol<-                                                                                      mapIds(edb,keys=rownames(res_contr_TBT_TC_anno),column="SYMBOL",keytype="GENEID",multiVals="first")
dim(res_contr_TBT_TC_anno) #25143 7 (just added a column for symbol)
head(res_contr_TBT_TC_anno)
library(writexl)
writexl::write_xlsx(res_contr_TBT_TC_anno,"~/Documents/Git/TBT_Paper_RNA_Work/Results/res_contr_TBT_TC_anno.xlsx")
```
##Subset significant results
```{r Significant normalized counts and volcano plots}
#All significant normalized counts table
res_contr_TBT_TC_padj_0.05 <- res_contr_TBT_TC_anno[which(res_contr_TBT_TC_anno$padj<0.05),]
dim(res_contr_TBT_TC_padj_0.05) #7920 DEGs over all when MSCs included
res_sig_order_TBT_TC <-res_contr_TBT_TC_padj_0.05[order(res_contr_TBT_TC_padj_0.05$padj),]
## Significant counts tale with normalized counts
res_sig_normalized_counts_TBT_TC <- assay(vsd_TBT_TC)[substr(rownames(res_sig_order_TBT_TC),1,15),] 
head(res_sig_normalized_counts_TBT_TC)


###Subsetting for upa nddown regullatd
##Significant in each wat tables
res_TBT_TC_sig_down <-res_contr_TBT_TC_padj_0.05[which(res_contr_TBT_TC_padj_0.05$log2FoldChange<0),] 
dim(res_TBT_TC_sig_down) #3764 genes
res_TBT_TC_sig_up <-res_contr_TBT_TC_padj_0.05[which(res_contr_TBT_TC_padj_0.05$log2FoldChange>0),]
dim(res_TBT_TC_sig_up) #4156 genes

#Normalized counts table for up and down regulations
res_sig_down_normalized_counts_TBT_TC <- res_sig_normalized_counts_TBT_TC[substr(rownames(res_TBT_TC_sig_down),1,15),]
res_sig_up_normalized_counts_TBT_TC <- res_sig_normalized_counts_TBT_TC[substr(rownames(res_TBT_TC_sig_up),1,15),]

##Save DEG tables for downstream analysis
library(writexl)
writexl::write_xlsx(res_sig_order_TBT_TC,"~/Documents/Git/TBT_Paper_RNA_Work/Results/res_sig_order_TBT_TC.xlsx")
```
#Case Control Removing MSC condition
```{r, eval=FALSE}

#BiocManager::install("DESeq2")
suppressPackageStartupMessages(library("DESeq2"))

#Subsetting the data to only have the pertinent information
cts_mat_TBT_No_MSC<- cts_mat[,which(substr(colnames(cts_mat),2,2)=="T"|substr(colnames(cts_mat),2,2)=="D")]

coldata_TBT_No_MSC <- coldata_norm[which(substr(rownames(coldata_norm),2,2)=="T"|substr(rownames(coldata_norm),2,2)=="D"),]
dim(cts_mat_TBT_No_MSC) #58037 28
dim(coldata_TBT_No_MSC) #28 4
coldata_TBT
##Creating DDS object from the subset data                                  

dds_TBT_NoMSC <- DESeqDataSetFromMatrix(countData = cts_mat_TBT_No_MSC,
                                                 colData = coldata_TBT_No_MSC,
                                                 design = ~ Day + condition
                                                 )

summary(dds_TBT_NoMSC)
head(dds_TBT_NoMSC)   
#####Adding features and metadata columns
rownames(dds_TBT_NoMSC) <- substr(rownames(dds_TBT_NoMSC),1,15)
featureData <- data.frame(ensgene= substr(rownames(cts_mat_TBT_No_MSC),1,15))
dim(featureData)
featureData$symbol <- cts_1021$SYMBOL# Adding mcol symbol data
mcols(dds_TBT_NoMSC) <- DataFrame(mcols(dds_TBT_NoMSC), featureData)
mcols(dds_TBT_NoMSC)

## Filtering DDS Object
#######removing low count rows #########
keep <- rowSums(counts(dds_TBT_NoMSC)) >= 50 #24643 genes passed this low 
dds_TBT_NoMSC <- dds_TBT_NoMSC[keep,]
dim(dds_TBT_NoMSC)#24643 28

#######Factor creation ##############
dds_TBT_NoMSC$condition <- factor(dds_TBT_NoMSC$condition, levels = c("control","TBT"))
dds_TBT_NoMSC$condition
dds_TBT_NoMSC$Day
```

### NORMALIZE: This is self-explanitory. This will normalize the data so
that we aren't getting a lot of outliers.
```{r normalize }
# Determine the size factors to use for normalization
dds_TBT_NoMSC<- estimateSizeFactors(dds_TBT_NoMSC)
sizeFactors(dds_TBT_NoMSC)
# Extract the normalized counts
normalized_counts_TBT_NoMSC <- counts(dds_TBT_NoMSC, normalized=T)
head(normalized_counts_TBT_NoMSC)
```
### VST vs rlog
VST is much faster ot compute and less sensitive to outliers than the rlog. rlog is better for smaller datasets with samples sizeds below ~30. you can perform both VST and rlog if you are unsure and then compare the meanSdPlot  or PCAs to see which generates better data. 
```{r correlation heatmaps}
# Transform the normalized counts 
vsd_TBT_NoMSC <- vst(dds_TBT_NoMSC,blind=F)
# Extract the matrix of transformed counts
vsd_mat_TBT_NoMSC <- assay(dds_TBT_NoMSC)
# Compute the correlation values between samples
vsd_cor_TBT_NoMSC <- cor(vsd_mat_TBT_NoMSC) 

# Plot the heatmap
pheatmap(vsd_cor_TBT_NoMSC)
colData(vsd_TBT_NoMSC)
###Plot PCAs####
plotPCA(vsd_TBT_NoMSC, intgroup= "condition")
plotPCA(vsd_TBT_NoMSC, intgroup = c("Time","condition"))
```
#Results from DESEQ2Diff expression analysis
```{r Run DESEQ DEG analysis and save data}
### RUN DESEQ2 for diffferential expression############
DES_TBT_NoMSC <- DESeq(dds_TBT_NoMSC)
res_TBT_NoMSC <- results(DES_TBT_NoMSC)
resultsNames(res_TBT_NoMSC)
head(res_TBT_NoMSC)
mcols(res_TBT_NoMSC)
#Increase Alpha threshold
res_TBT_NoMSC_contrast <- results(DES_TBT_NoMSC, contrast=c("condition","TBT","control"),alpha = 0.05)
head(res_TBT_NoMSC_contrast)

########Annotate significant genes#################
library(AnnotationDbi)
library(org.Hs.eg.db)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(EnsDb.Hsapiens.v86)
edb <- EnsDb.Hsapiens.v86
organism(edb)
keytypes(edb)
#keys(edb,keytype = "GENEID")

## Annotating analyzed results the data can be subset for analysis and significance later
dim(res_TBT_NoMSC_contrast)#24643 6
res_TBT_NoMSC_contrast_anno <- data.frame(res_TBT_NoMSC_contrast)
res_TBT_NoMSC_contrast_anno$symbol<-                                                                                      mapIds(edb,keys=rownames(res_TBT_NoMSC_contrast_anno),column="SYMBOL",keytype="GENEID",multiVals="first")
dim(res_TBT_NoMSC_contrast_anno) #24643 7 (just added a column for symbol)
head(res_TBT_NoMSC_contrast_anno)
## Save data
library(writexl)
writexl::write_xlsx(res_TBT_NoMSC_contrast_anno,"~/Documents/Git/TBT_Paper_RNA_Work/Results/res_TBT_NoMSC_contrast_anno.xlsx")
#####Subsetting for significance######3
res_TBT_NoMSC_contrast_padj_0.05 <- res_TBT_NoMSC_contrast_anno[which(res_TBT_NoMSC_contrast_anno$padj<0.05),]


#All significant normalized counts table
res_order_TBT_NoMSC <-res_TBT_NoMSC_contrast_padj_0.05[order(res_TBT_NoMSC_contrast_padj_0.05$padj),]

res_sig_order_TBTcontrNoMSC_padj_0.05 <- res_TBT_NoMSC_contrast_padj_0.05[order(res_TBT_NoMSC_contrast_padj_0.05$padj),]
dim(res_order_TBT_NoMSC)#7686 gene Differentially expressed
###Save Data table
library(writexl)
writexl::write_xlsx(res_order_TBT_NoMSC,"~/Documents/Git/TBT_Paper_RNA_Work/Results/res_order_TBT_NoMSC.xlsx")
```


#TBT Day vs Day comparisons
This will be a simple analysis using the condition as the design element. 
#D3 Analysis
```{r TBT Day 3 comparison}
#BiocManager::install("DESeq2")
suppressPackageStartupMessages(library("DESeq2"))

#Subsetting the data to only have the pertinent information
# cts_mat_TBT<- cts_mat[,which(substr(colnames(cts_mat),2,2)=="T"|substr(colnames(cts_mat),2,2)=="D")]
# rownames(cts_mat_TBT) <- substr(rownames(cts_mat_TBT),1,15)#Removing the .# on the gene names
# coldata_TBT <- coldata_norm[which(substr(rownames(coldata_norm),2,2)=="T"|substr(rownames(coldata_norm),2,2)=="D"),]
# dim(cts_mat_TBT) #58037 27
colnames(cts_mat_TBT_No_MSC)
cts_mat_TBT_D3 <- cts_mat_TBT_No_MSC[,1:7]
coldata_TBT_D3 <- coldata_TBT_No_MSC[1:7,]
##Creating DDS object from the subset data                                  
dds_TBT_D3 <- DESeqDataSetFromMatrix(countData = cts_mat_TBT_D3,
                                                 colData = coldata_TBT_D3,
                                                 design = ~ condition                                               )

summary(dds_TBT_D3)
head(dds_TBT_D3)   
###Adding meta data columns
rownames(dds_TBT_D3) <- substr(rownames(dds_TBT_D3),1,15)
featureData <- data.frame(gene=substr(rownames(cts_mat_TBT_D3),1,15))
mcols(dds_TBT_D3) <- DataFrame(mcols(dds_TBT_D3), featureData)
mcols(dds_TBT_D3)
featureData$symbol <- cts_1021$SYMBOL# Adding mcol symbol data
mcols(dds_TBT_D3) <- DataFrame(mcols(dds_TBT_D3), featureData)
mcols(dds_TBT_D3)
###Removing low quality rows
keep <- rowSums(counts(dds_TBT_D3)) >= 50
length(keep) #58037
dds_TBT_D3 <- dds_TBT_D3[keep,]
####Making things factors
dds_TBT_D3$condition <- factor(dds_TBT_D3$condition, levels = c("control","TBT"))
dds_TBT_D3$condition
# Determine the size factors to use for normalization
dds_TBT_D3<- estimateSizeFactors(dds_TBT_D3)
sizeFactors(dds_TBT_D3)
# Extract the normalized counts
normalized_counts_TBT_D3 <- counts(dds_TBT_D3, normalized=T)
head(normalized_counts_TBT_D3)

# Transform the normalized counts 
vsd_TBT_D3 <- vst(dds_TBT_D3,blind=F )

# Extract the matrix of transformed counts
vsd_mat_TBT_D3 <- assay(dds_TBT_D3)

# Compute the correlation values between samples
vsd_cor_TBT_D3 <- cor(vsd_mat_TBT_D3) 

# Plot the heatmap
pheatmap(vsd_cor_TBT_D3)

colData(vsd_TBT_D3)
###Plot PCAs####
plotPCA(vsd_TBT_D3, intgroup= "condition")

plotPCA(vsd_TBT_D3, intgroup = c("Time","condition"))

DES_TBT_D3 <- DESeq(dds_TBT_D3)
res_TBT_D3 <- results(DES_TBT_D3)
resultsNames(res_TBT_D3)
head(res_TBT_D3)
mcols(res_TBT_D3)
#Increase Alpha threshold
res_TBT_D3_contrast <- results(DES_TBT_D3, contrast=c("condition","TBT","control"),alpha = 0.05)
head(res_TBT_D3_contrast)

##Annnotate DEG analysis table
suppressPackageStartupMessages(library(org.Hs.eg.db))
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(EnsDb.Hsapiens.v86)
edb <- EnsDb.Hsapiens.v86
organism(edb)
keytypes(edb)
###Annotate results table#####
res_TBT_D3_contrast_anno <- data.frame(res_TBT_D3_contrast)
res_TBT_D3_contrast_anno$symbol <- mapIds(edb,keys=rownames(res_TBT_D3_contrast_anno),column="SYMBOL",keytype="GENEID",multiVals="first")
## Save data
library(writexl)
writexl::write_xlsx(res_TBT_D3_contrast_anno,"~/Documents/Git/TBT_Paper_RNA_Work/Results/res_TBT_D3_contrast_anno.xlsx")

### Subsetting for padj data ####
sum(res_TBT_D3_contrast_anno$padj < 0.05, na.rm=TRUE) #2782 lowering the Padjusted value removed some of the genes
res_TBT_D3_contrast_padj_0.05 <- res_TBT_D3_contrast_anno[which(res_TBT_D3_contrast_anno$padj<0.05),]


#All significant normalized counts table
res_order_TBT_D3 <-res_TBT_D3_contrast_padj_0.05[order(res_TBT_D3_contrast_padj_0.05$padj),]

dim(res_order_TBT_D3)#2782 6
## Save tables
library(writexl)
writexl::write_xlsx(res_order_TBT_D3,"~/Documents/Git/MSC_Bulk/RNA_seq_Data/TBT_results/res_order_TBT_D3.xlsx")
```

###D7
```{r Day 7 comparison}
#BiocManager::install("DESeq2")
suppressPackageStartupMessages(library("DESeq2"))

#Subsetting the data to only have the pertinent information
# cts_mat_TBT<- cts_mat[,which(substr(colnames(cts_mat),2,2)=="R"|substr(colnames(cts_mat),2,2)=="D")]
# rownames(cts_mat_TBT) <- substr(rownames(cts_mat_TBT),1,15)#Removing the .# on the gene names
# coldata_TBT <- coldata_norm[which(substr(rownames(coldata_norm),2,2)=="R"|substr(rownames(coldata_norm),2,2)=="D"),]
colnames(cts_mat_TBT_No_MSC) #58037 27
# dim(coldata_TBT) #27 4
cts_mat_TBT_D7 <- cts_mat_TBT_No_MSC[,8:15]
coldata_TBT_D7 <- coldata_TBT_No_MSC[8:15,]
##Creating DDS object from the subset data     
head(cts_mat_TBT_D7)
dds_TBT_D7 <- DESeqDataSetFromMatrix(countData = cts_mat_TBT_D7,
                                                 colData = coldata_TBT_D7,
                                                 design = ~ condition                                               )
summary(dds_TBT_D7)
head(dds_TBT_D7)   
###Adding meta data columns
rownames(dds_TBT_D7) <- substr(rownames(dds_TBT_D7),1,15)
featureData <- data.frame(gene=substr(rownames(cts_mat_TBT_D7),1,15))
mcols(dds_TBT_D7) <- DataFrame(mcols(dds_TBT_D7), featureData)
mcols(dds_TBT_D7)
featureData$symbol <- cts_1021$SYMBOL# Adding mcol symbol data
mcols(dds_TBT_D7) <- DataFrame(mcols(dds_TBT_D7), featureData)
mcols(dds_TBT_D7)
###Removing low quality rows
keep <- rowSums(counts(dds_TBT_D7)) >= 50
length(keep) #24448 
dds_TBT_D7 <- dds_TBT_D7[keep,]
####Making things factors
dds_TBT_D7$condition <- factor(dds_TBT_D7$condition, levels = c("control","TBT"))
dds_TBT_D7$condition
# Determine the size factors to use for normalization
dds_TBT_D7<- estimateSizeFactors(dds_TBT_D7)
sizeFactors(dds_TBT_D7)
# Extract the normalized counts
normalized_counts_TBT_D7 <- counts(dds_TBT_D7, normalized=T)
head(normalized_counts_TBT_D7)

# Transform the normalized counts 
vsd_TBT_D7 <- vst(dds_TBT_D7,blind=F )

# Extract the matrix of transformed counts
vsd_mat_TBT_D7 <- assay(dds_TBT_D7)

# Compute the correlation values between samples
vsd_cor_TBT_D7 <- cor(vsd_mat_TBT_D7) 

# Plot the heatmap
pheatmap(vsd_cor_TBT_D7)

colData(vsd_TBT_D7)
###Plot PCAs####
plotPCA(vsd_TBT_D7, intgroup= "condition")

plotPCA(vsd_TBT_D7, intgroup = c("Time","condition"))

DES_TBT_D7 <- DESeq(dds_TBT_D7)
res_TBT_D7 <- results(DES_TBT_D7)
resultsNames(res_TBT_D7)
head(res_TBT_D7)
mcols(res_TBT_D7)
#Increase Alpha threshold
res_TBT_D7_contrast <- results(DES_TBT_D7, contrast=c("condition","TBT","control"),alpha = 0.05)
head(res_TBT_D7_contrast)

##Annnotate DEG analysis table
suppressPackageStartupMessages(library(org.Hs.eg.db))
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(EnsDb.Hsapiens.v86)
edb <- EnsDb.Hsapiens.v86
organism(edb)
keytypes(edb)
###Annotate results table#####
res_TBT_D7_contrast_anno <- data.frame(res_TBT_D7_contrast)
res_TBT_D7_contrast_anno$symbol <- mapIds(edb,keys=rownames(res_TBT_D7_contrast_anno),column="SYMBOL",keytype="GENEID",multiVals="first")
## Save data
library(writexl)
writexl::write_xlsx(res_TBT_D7_contrast_anno,"~/Documents/Git/TBT_Paper_RNA_Work/Results/res_TBT_D7_contrast_anno.xlsx")

### Subsetting for padj data ####
sum(res_TBT_D7_contrast_anno$padj < 0.05, na.rm=TRUE) #6267 lowering the Padjusted value removed some of the genes
res_TBT_D7_contrast_padj_0.05 <- res_TBT_D7_contrast_anno[which(res_TBT_D7_contrast_anno$padj<0.05),]


#All significant normalized counts table
res_order_TBT_D7 <-res_TBT_D7_contrast_padj_0.05[order(res_TBT_D7_contrast_padj_0.05$padj),]

dim(res_order_TBT_D7)#6267 6
## Save tables
library(writexl)
writexl::write_xlsx(res_order_TBT_D7,"~/Documents/Git/MSC_Bulk/RNA_seq_Data/TBT_results/res_order_TBT_D7.xlsx")
```

###D10
```{r Day 10 comparison}
#BiocManager::install("DESeq2")
suppressPackageStartupMessages(library("DESeq2"))

#Subsetting the data to only have the pertinent information
# cts_mat_TBT<- cts_mat[,which(substr(colnames(cts_mat),2,2)=="T"|substr(colnames(cts_mat),2,2)=="D")]
# rownames(cts_mat_TBT) <- substr(rownames(cts_mat_TBT),1,15)#Removing the .# on the gene names
# coldata_TBT <- coldata_norm[which(substr(rownames(coldata_norm),2,2)=="T"|substr(rownames(coldata_norm),2,2)=="D"),]
colnames(cts_mat_TBT_No_MSC) #58037 27
# dim(coldata_TBT) #27 4
cts_mat_TBT_D10 <- cts_mat_TBT_No_MSC[,16:22]
coldata_TBT_D10 <- coldata_TBT_No_MSC[16:22,]
##Creating DDS object from the subset data                                  
dds_TBT_D10 <- DESeqDataSetFromMatrix(countData = cts_mat_TBT_D10,
                                                 colData = coldata_TBT_D10,
                                                 design = ~ condition                                               )
summary(dds_TBT_D10)
head(dds_TBT_D10)   
###Adding meta data columns
rownames(dds_TBT_D10) <- substr(rownames(dds_TBT_D10),1,15)
featureData <- data.frame(gene=substr(rownames(cts_mat_TBT_D10),1,15))
mcols(dds_TBT_D10) <- DataFrame(mcols(dds_TBT_D10), featureData)
mcols(dds_TBT_D10)
featureData$symbol <- cts_1021$SYMBOL# Adding mcol symbol data
mcols(dds_TBT_D10) <- DataFrame(mcols(dds_TBT_D10), featureData)
mcols(dds_TBT_D10)
###Removing low quality rows
keep <- rowSums(counts(dds_TBT_D10)) >= 50
length(keep) #58037
dds_TBT_D10 <- dds_TBT_D10[keep,]
####Making things factors
dds_TBT_D10$condition <- factor(dds_TBT_D10$condition, levels = c("control","TBT"))
dds_TBT_D10$condition
# Determine the size factors to use for normalization
dds_TBT_D10<- estimateSizeFactors(dds_TBT_D10)
sizeFactors(dds_TBT_D10)
# Extract the normalized counts
normalized_counts_TBT_D10 <- counts(dds_TBT_D10, normalized=T)
head(normalized_counts_TBT_D10)

# Transform the normalized counts 
vsd_TBT_D10 <- vst(dds_TBT_D10,blind=F )

# Extract the matrix of transformed counts
vsd_mat_TBT_D10 <- assay(dds_TBT_D10)

# Compute the correlation values between samples
vsd_cor_TBT_D10 <- cor(vsd_mat_TBT_D10) 

# Plot the heatmap
pheatmap(vsd_cor_TBT_D10)

colData(vsd_TBT_D10)
###Plot PCAs####
plotPCA(vsd_TBT_D10, intgroup= "condition")

plotPCA(vsd_TBT_D10, intgroup = c("Time","condition"))
######Results DESEQ############
DES_TBT_D10 <- DESeq(dds_TBT_D10)
res_TBT_D10 <- results(DES_TBT_D10)
resultsNames(res_TBT_D10)
head(res_TBT_D10)
mcols(res_TBT_D10)
#Increase Alpha threshold
res_TBT_D10_contrast <- results(DES_TBT_D10, contrast=c("condition","TBT","control"),alpha = 0.05)
head(res_TBT_D10_contrast)

##Annnotate DEG analysis table
suppressPackageStartupMessages(library(org.Hs.eg.db))
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(EnsDb.Hsapiens.v86)
edb <- EnsDb.Hsapiens.v86
organism(edb)
keytypes(edb)
###Annotate results table#####
res_TBT_D10_contrast_anno <- data.frame(res_TBT_D10_contrast)
res_TBT_D10_contrast_anno$symbol <- mapIds(edb,keys=rownames(res_TBT_D10_contrast_anno),column="SYMBOL",keytype="GENEID",multiVals="first")
## Save data
library(writexl)
writexl::write_xlsx(res_TBT_D10_contrast_anno,"~/Documents/Git/TBT_Paper_RNA_Work/Results/res_TBT_D10_contrast_anno.xlsx")

### Subsetting for padj data ####
sum(res_TBT_D10_contrast_anno$padj < 0.05, na.rm=TRUE) #4424 lowering the Padjusted value removed some of the genes
res_TBT_D10_contrast_padj_0.05 <- res_TBT_D10_contrast_anno[which(res_TBT_D10_contrast_anno$padj<0.05),]


#All significant normalized counts table
res_order_TBT_D10 <-res_TBT_D10_contrast_padj_0.05[order(res_TBT_D10_contrast_padj_0.05$padj),]

dim(res_order_TBT_D10)#4424 6
## Save tables
library(writexl)
writexl::write_xlsx(res_order_TBT_D10,"~/Documents/Git/MSC_Bulk/RNA_seq_Data/TBT_results/res_order_TBT_D10.xlsx")
```

###D14
```{r Day 14 comparison}
#BiocManager::install("DESeq2")
suppressPackageStartupMessages(library("DESeq2"))

#Subsetting the data to only have the pertinent information
# cts_mat_TBT<- cts_mat[,which(substr(colnames(cts_mat),2,2)=="R"|substr(colnames(cts_mat),2,2)=="D")]
# rownames(cts_mat_TBT) <- substr(rownames(cts_mat_TBT),1,15)#Removing the .# on the gene names
# coldata_TBT <- coldata_norm[which(substr(rownames(coldata_norm),2,2)=="R"|substr(rownames(coldata_norm),2,2)=="D"),]
colnames(cts_mat_TBT) #58037 27
# dim(coldata_TBT) #27 4
cts_mat_TBT_D14 <- cts_mat_TBT_No_MSC[,23:28]
coldata_TBT_D14 <- coldata_TBT_No_MSC[23:28,]
##Creating DDS object from the subset data                                  
dds_TBT_D14 <- DESeqDataSetFromMatrix(countData = cts_mat_TBT_D14,
                                                 colData = coldata_TBT_D14,
                                                 design = ~ condition                                               )
summary(dds_TBT_D14)
head(dds_TBT_D14)   
###Adding meta data columns
rownames(dds_TBT_D14) <- substr(rownames(dds_TBT_D14),1,15)
featureData <- data.frame(gene=substr(rownames(cts_mat_TBT_D14),1,15))
mcols(dds_TBT_D14) <- DataFrame(mcols(dds_TBT_D14), featureData)
mcols(dds_TBT_D14)
featureData$symbol <- cts_1021$SYMBOL# Adding mcol symbol data
mcols(dds_TBT_D14) <- DataFrame(mcols(dds_TBT_D14), featureData)
mcols(dds_TBT_D14)
###Removing low quality rows
keep <- rowSums(counts(dds_TBT_D14)) >= 50
length(keep) #24448 
dds_TBT_D14 <- dds_TBT_D14[keep,]
####Making things factors
dds_TBT_D14$condition <- factor(dds_TBT_D14$condition, levels = c("control","TBT"))
dds_TBT_D14$condition
# Determine the size factors to use for normalization
dds_TBT_D14<- estimateSizeFactors(dds_TBT_D14)
sizeFactors(dds_TBT_D14)
# Extract the normalized counts
normalized_counts_TBT_D14 <- counts(dds_TBT_D14, normalized=T)
head(normalized_counts_TBT_D14)

# Transform the normalized counts 
vsd_TBT_D14 <- vst(dds_TBT_D14,blind=F )

# Extract the matrix of transformed counts
vsd_mat_TBT_D14 <- assay(dds_TBT_D14)

# Compute the correlation values between samples
vsd_cor_TBT_D14 <- cor(vsd_mat_TBT_D14) 

# Plot the heatmap
pheatmap(vsd_cor_TBT_D14)

colData(vsd_TBT_D14)
###Plot PCAs####
plotPCA(vsd_TBT_D14, intgroup= "condition")

plotPCA(vsd_TBT_D14, intgroup = c("Time","condition"))

DES_TBT_D14 <- DESeq(dds_TBT_D14)
res_TBT_D14 <- results(DES_TBT_D14)
resultsNames(res_TBT_D14)
head(res_TBT_D14)
mcols(res_TBT_D14)
#Increase Alpha threshold
res_TBT_D14_contrast <- results(DES_TBT_D14, contrast=c("condition","TBT","control"),alpha = 0.05)
head(res_TBT_D14_contrast)

##Annnotate DEG analysis table
suppressPackageStartupMessages(library(org.Hs.eg.db))
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(EnsDb.Hsapiens.v86)
edb <- EnsDb.Hsapiens.v86
organism(edb)
keytypes(edb)
###Annotate results table#####
res_TBT_D14_contrast_anno <- data.frame(res_TBT_D14_contrast)
res_TBT_D14_contrast_anno$symbol <- mapIds(edb,keys=rownames(res_TBT_D14_contrast_anno),column="SYMBOL",keytype="GENEID",multiVals="first")
## Save data
library(writexl)
writexl::write_xlsx(res_TBT_D14_contrast_anno,"~/Documents/Git/TBT_Paper_RNA_Work/Results/res_TBT_D14_contrast_anno.xlsx")

### Subsetting for padj data ####
sum(res_TBT_D14_contrast_anno$padj < 0.05, na.rm=TRUE) #4836 lowering the Padjusted value removed some of the genes
res_TBT_D14_contrast_padj_0.05 <- res_TBT_D14_contrast_anno[which(res_TBT_D14_contrast_anno$padj<0.05),]


#All significant normalized counts table
res_order_TBT_D14 <-res_TBT_D14_contrast_padj_0.05[order(res_TBT_D14_contrast_padj_0.05$padj),]

dim(res_order_TBT_D14)#4836 6
## Save tables
library(writexl)
writexl::write_xlsx(res_order_TBT_D14,"~/Documents/Git/MSC_Bulk/RNA_seq_Data/TBT_results/res_order_TBT_D14.xlsx")
```


```{r Subset data for significance}
library(writexl)
#writexl::write_xlsx(res_TBT_contrNoMSC_anno,"~/Documents/Git/MSC_Bulk/RNA_seq_Data/TBT_results/TBT_noMSC_DEGs_NoMSC_sig_anno.xlsx")
```